Title

### 2024-11-23 paper/paper.qmd
appendix for postrior done paper pdf render

analysis_data <- read_parquet(here::here("data/02-analysis_data/loblaws_data.parquet"))
```

@@

sorted_data <- analysis_data %>%
  arrange(date)

@@

filtered_data <- analysis_data %>%
  filter(product_id %in% selected_ids) %>%
  mutate(label = paste(product_name, size, sep = ", "))

@@

p2 <- ggplot(analysis_data, aes(x = size, y = current_price, fill = size)) +
  geom_boxplot() +
  labs(x = "Size", y = "Price", title = "Price Distribution by Size")  +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 5, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
)

@@

```{r, message=FALSE, warning=FALSE, fig.width= 7, fig.height= 3}
#| label: fig-2
#| fig-cap: Average Support Rate  for Donald Trump from 2024-07-21 to 2024-10-23 reported by different pollsters
#| echo: false
# Group data by date and product type, and calculate average price for each type per date
type_trend_data <- analysis_data %>%
  group_by(date, type) %>%
  summarise(avg_price = mean(current_price, na.rm = TRUE)) %>%
  ungroup()

@@

# Group data by date and product size, and calculate average price for each size per date
size_trend_data <- analysis_data %>%
  group_by(date, size) %>%
  summarise(avg_price = mean(current_price, na.rm = TRUE)) %>%
  ungroup()

@@

```{r, message=FALSE}
#| echo: false
#| eval: false
#| warning: false
#| message: false

@@

>>>>

loblaws_data <- read_parquet(here::here("data/02-analysis_data/loblaws_data.parquet"))
```

@@

sorted_data <- loblaws_data %>%
  arrange(date)

@@

filtered_data <- loblaws_data %>%
  filter(product_id %in% selected_ids) %>%
  mutate(label = paste(product_name, size, sep = ", "))

@@

p2 <- ggplot(loblaws_data, aes(x = size, y = current_price, fill = size)) +
  geom_boxplot() +
  labs(x = "Size", y = "Price", title = "Price Distribution by Size")  +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    plot.title = element_text(size = 7, face = "bold", color = "#333333"),
    plot.subtitle = element_text(size = 6, color = "#333333"),
    legend.title = element_text(size = 5, color = "#333333"),
    legend.text = element_text(size = 5, color = "#333333"),
    legend.key.size = unit(0.5, "lines"),
    axis.title = element_text(size = 6, color = "#333333"),
    strip.text = element_text(size = 5)
)

@@

```{r, message=FALSE, warning=FALSE, fig.width= 7, fig.height= 3}
#| label: fig-2
#| fig-cap: Average Support Rate  for Donald Trump from 2024-07-21 to 2024-10-23 reported by different pollsters
#| echo: false
# Group data by date and product type, and calculate average price for each type per date
type_trend_data <- loblaws_data %>%
  group_by(date, type) %>%
  summarise(avg_price = mean(current_price, na.rm = TRUE)) %>%
  ungroup()

@@

# Group data by date and product size, and calculate average price for each size per date
size_trend_data <- loblaws_data %>%
  group_by(date, size) %>%
  summarise(avg_price = mean(current_price, na.rm = TRUE)) %>%
  ungroup()

@@

@@

\appendix

# Appendix {-}

# Model details {#sec-model-details}
R package `bayesplot` [@bayesplot] is used for posteriors.

Due to space limitations, we have chosen to display only the posterior and prior comparisons for a subset of variables in the posterior analysis. If we were to display all variables included in the model, the chart would become overly complex and difficult to interpret. Although the effects of multiple variables differ, focusing on this subset may give the impression that the model’s sensitivity to variations across different factors is limited. Thus, we have selected representative variables, including the intercept, size categories ("sizeLarge" and "sizeSmall"), and product types ("typeSnacks," "typeFlour/Rice," and "typeSeasoning"). The output results for other variables not shown are generally consistent with the trends observed here. Additionally, in regression models (such as stan_glmer), categorical variables (factors) are typically automatically converted into dummy variables. One category is set as the reference category (baseline), and its effect is included in the intercept, while the coefficients for other categories represent their offsets relative to the baseline. For example, in our model, if `Extra Large` is set as the baseline (default), it will not appear as a separate coefficient. The coefficients for `sizeLarge` and `sizeSmall` would then represent their deviations relative to `Extra Large`.

## Posterior predictive check
@fig-post1 displays a comparison between the distribution of observed data $y$ and the distribution of simulated data $y_{rep}$ generated by our model. The dark solid line represents the observed data's actual distribution, while the multiple light lines correspond to distributions from various simulated datasets. If the distribution of simulated data $y_{rep}$ closely resembles that of the observed data $y$, it suggests that the model effectively captures the underlying characteristics of the observed data.

In @fig-post1, we observe that the simulated data distributions (light lines) align closely with the real data distribution (dark line) across the entire range, particularly at the peaks and troughs of the observed data. This indicates that the model has captured the distributional features of the observed data, including its skewness and multimodal structure. The ability to replicate these features demonstrates the model's strong predictive capacity and its reliability in estimating posterior distributions. This result builds confidence in the model's suitability for further inference and prediction.

```{r, fig.width=5, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-post1
#| fig-cap: PP Check for Model
#| warning: false

pp_check(first_model)
```

@fig-post2 displays the MCMC sampling trace plots for six key parameters in our model, including the `(Intercept)`, size variables `sizeLarge` and ` sizeSmall`, and type variables `typeSnacks`, `typeFlour/Rice` and `typeSeasoning`. Each plot includes four chains from Chain 1 to Chain 4, representing independent MCMC sampling processes. In a well-converged model, our expectation is that each chain to stabilize and fluctuate around a central value without showing a clear trend (i.e., the mean of the samples converges over time). Additionally, the chains for a parameter should exhibit consistent behavior, indicating agreement among the chains.

From @fig-post2, we observe that the sampling chains for all six parameters show a steady state, with fluctuations around a central value and no observable trends over the iterations. Moreover, the chains for each parameter are closely aligned, demonstrating consistency between the chains. This isuggests that the model has reached convergence and the posterior distributions for these parameters are stable. These trace plots support the reliability of our MCMC sampling process and provide confidence in the validity of the posterior estimates.

```{r, fig.width=7, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-post2
#| fig-cap: MCMC sampling trace plots for model
#| warning: false

plot(first_model, "trace", pars = c("(Intercept)", "sizeLarge", "sizeSmall", "typeSnacks", "typeFlour/Rice", "typeSeasoning"))
```
@fig-post3 shows the $\hat{R}$ diagnostic statistic for our model, which is used to assess the convergence of MCMC sampling. Ideally, is that the $\hat{R}$ value would be close to value 1, which indicates consistent variance across different MCMC chains, signifying that the model has reached a stable state. The horizontal axis represents the $\hat{R}$ values, where values closer to 1 indicate better convergence. Our expectation is that, all parameters would have $\hat{R}$ values below 1.05, which means minimal variance differences between chains and consistency across them. This plot shows that all $\hat{R}$ values are concentrated around 1, and all $\hat{R}$ values are lower than the threshold of 1.05 (indicated by the vertical dashed line on the right), suggesting that the sampling chains have converged well, providing thw reliability for our model results.

```{r, fig.width=5, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-post3
#| fig-cap: R hat Diagnostic statistic for model
#| warning: false

plot(first_model, "rhat")
```

@fig-post4 shows the posterior distributions of certain parameters in our model, including `(Intercept)`, size variables `sizeLarge` and `sizeSmall`, and type variables `typeSnacks`, `typeFlour/Rice`, and `typeSeasoning`. The shape and location of these distributions help us understand the extent to which the observed data supports these parameter estimates and the degree of uncertainty associated with them.

The `(Intercept)` is centered nearly around 2.5, reflecting the baseline level for product prices in our model when all other factors are held constant. The narrow and peaked distribution suggests strong support for this estimate with relatively low uncertainty. For `sizeLarge` and `sizeSmall`, the distributions are shifted slightly to the left, indicating that these sizes are generally associated with lower prices compared to the baseline category `ExtraLarge`. Type variables `typeSnacks`, `typeFlour/Rice`, and `typeSeasoning` exhibit posterior distributions centered close to zero, implying that their influence on product pricing is minimal or uncertain, as the data does not strongly support a directional impact.

Overall, the results suggest that the model successfully captures the baseline pricing level and the relative influence of product sizes, but it struggles to discern clear effects for product types, possibly due to overlapping pricing patterns or insufficient variation within these categories.

```{r, fig.width=5, fig.height=3}
#| echo: false
#| eval: true
#| label: fig-post4
#| fig-cap: posterior distributions of four parameters in model
#| warning: false

plot(first_model, "areas", pars = c("(Intercept)", "sizeLarge", "sizeSmall", "typeSnacks", "typeFlour/Rice", "typeSeasoning"))
```
